---
title: "Assessing Relationship Among Brain Regions in Mice"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
  pdf_document: default
---

<style type="text/css">
body{ /* Normal  */
      font-size: 18px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```


Team ID: 5

Name (tasks): Oscar Alvardo

Name (tasks): Karshima Johnson

Name (tasks): Christina Thai

Auditor (tasks):

***

# Introduction

This report seeks to examine whether their exists a regulatory relationship among regions. In other words, it will investigate if activity in one region affects the activity in another region. 


## Background
```{r, include=FALSE}
library("R.matlab") # Install 'R.matlab' if this is the first time you call it.

ephysroot = '/Users/christinathai/Desktop/R_programming/STA_141A/Project/'; # path of the data set 
mstr = 'Krebs'; # mouse names

### Load saved data:
exp_data<- readMat(paste(ephysroot,mstr,'_reduced.mat',sep=''));
# check out data structure
str(exp_data)
```

<span style="color:red">Note: </span>  Discuss the basic background of this project here, even if you pick one of the three provided data sets. In particular, you are expected to briefly comment on the importance of this analysis, and explain the goals of this project. For instance, you can write two short paragraphs as follows.

In this project, we analyze the data set given ____ containing Neuropixel recordings of a mouse named Krebs during spontaneous activity. The data used has already been preprocessed using a Matlab script, which can be found in the code Appendix of the report. 

A brief introduction of cortical responses to sensory stimuli is as follows:

***
Cortical responses to sensory stimuli are highly variable, and sensory cortex exhibits intricate spontaneous activity even without external sensory input. Cortical variability and spontaneous activity have been variously proposed to represent random noise, recall of prior experience, or encoding of ongoing behavioral and cognitive variables. Here, by recording over 10,000 neurons in mouse visual cortex, we show that spontaneous activity reliably encodes a high-dimensional latent state, which is partially related to the ongoing behavior of the mouse and is represented not just in visual cortex but across the forebrain. Sensory inputs do not interrupt this ongoing signal, but add onto it a representation of visual stimuli in orthogonal dimensions. Thus, visual cortical population activity, despite its apparently noisy structure, reliably encodes an orthogonal fusion of sensory and multidimensional behavioral information.
***




`faceSVD`
`motionMask`

The nine regions for Kreubs in the mouse brain are:
`stall.CP` Caudate Putamen
`stall.FrMoCtx` Frontal Motor Cortex
`stall.HPF` Hippocampal Formation
`stall.LS`Lateral Septum
`stall.MB` Midbrain
`stall.SC` Superior Colliculus
`stall.SomMoCtxx` Somatomotor Cortex
`stall.TH` Thalamus
`stall.V1` Primary Visual Cortex

The data set contains spike trains of neurons in nine regions in the mouse brain

```{r}

brain_regions= ls(exp_data)[-c(1:3)]; 
n_time= dim(exp_data$faceSVD)[2];
n_sv = dim(exp_data$faceSVD)[1];
```





```{r}
i_region = 1; # pick a region
time_range = c(100, 500); # Pick a time window (indices of time frame)

### Extract the spike train for this region
spikes_this_region = exp_data[[brain_regions[1]]]

### Visualize the spike train 
n_neurons = dim(spikes_this_region)[1]; 
plot(x=0,y=0,pch=16,col="white",type="l",lwd=3,ylim=c(0,n_neurons),xlim=time_range,cex.lab=1,cex.main=1,ylab='Neuron',xlab='Time frame',main='Spikes') #yaxt='n',xaxt='n',

for(i_neuron in 1:n_neurons){
  spk_times =which(spikes_this_region[i_neuron,time_range[1]:time_range[2]]>0);
  if (length(spk_times)>0){
    points(y=rep(i_neuron,length(spk_times)),x=spk_times+time_range[1]-1,col="#000000",pch='.',cex=2)
  }
}

```




## Statistical questions of interest


<span style="color:red">Note: </span>  Translate the scientific goals into statistical questions that can be answered with appropriate statistical analysis on the given data.

To answer the primary scientific question of interest, 


# Analysis Plan


<span style="color:red">Note: </span>  It is best practice to write down the statistical analysis plan before analyzing data. This can significantly improve the rigorousity and reproducibility of the analysis, and avoid false discovery. This section should just be an extended version of your initial proposal. It is often inevitable that one may need to alter the analysis plan for unexpected features in the data. However, detailed documentation of the changes will shed light on unique features of the data set and help future analysis.


## Population and study design


<span style="color:red">Note: </span>  Discuss the target population of this analysis, and how the study is design to collect representative data for this population.

## Statistical Analysis

### Descriptive Analysis


<span style="color:red">Note: </span>  Describe the descriptive analysis you plan to conduct on the data set. The goal of the descriptive analysis is to provide an overview of the data set (and the study that collects the data) beforing diving into the analysis.

We will summarize the distribution of each variables for patients in the treatment group and the control group. Based on the summary statistics, we will evaluate whether the randomization is successful. To be specific, all summary statistics of baseline variables should be similar between the treatment and control groups.

We will estimate the survival function of patients in the treatment and control groups using Kaplan-Meier estimates, without adjusting for any covariates. Furthermore, we will summarize the distribution of observation time and the patients' status at the end of study in groups defined by treatment, regions of patients, indicator that patient had an abnormal LDF level, or indicator that patient had an abnormal alkaline phoshatase level.

### "Main" Analysis


<span style="color:red">Note: </span>  Describe the analysis you plan to conduct to answer the questions of interest. Remember to rename the section title to fit the actual analysis you propose, e.g., inferential analysis, clustering, prediction model, etc. For instance, in this report, this section should explain

* Detailed model specification of linear regression model.
* Description of the Cox proportional hazard model.
* Description of the model with interactions.

Furthermore, this section should also explain how you plan to interpret the conclusion---this is a convention for the analysis of clinical trials, which may not apply to other data science projects.



### Sensitivity Analysis


<span style="color:red">Note: </span>  Describe additional analysis you plan to conduct to complement the main analysis. This section should include analysis to show that the results you get are not sensitive to the model choices or assumptions. You can also conduct analysis to account for features in the data you have not foreseen.  


# Results

<span style="color:red">Note: </span> In this template, we present the unpolished tables and figures generated with generic `r` functions, and we skip all interpretation of the analysis results and the sensitivity analysis.  You can adapt any functions in this section in your report.

## Descriptive Analysis


<span style="color:red">Note: </span> There are many ways to automatically generate summary tables in `R`, of which many would even generate html or latex tables. In this template, we will use the package `qwraps2` as an example ([link](https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html)).

```{r, eval = FALSE, include = FALSE}
library(dplyr)
library(tidyr)
library(qwraps2)

options(digits = 3)  

set.seed(1) # Read more about RNGs with ?set.seed

### Read data
nsclc_data <- read.table("../Data/nsclc-modified.txt", header = TRUE)

excluded_vars = c("ptid","tx");
table1_summary <-  nsclc_data %>%   select(.,-one_of(excluded_vars)) %>%  qsummary(.)

table1 =nsclc_data %>% group_by(.data$tx) %>%
  summary_table(., table1_summary)

rgroups =   rep(4,length(names(nsclc_data))-2)
names(rgroups)=names(nsclc_data)[!names(nsclc_data)%in%excluded_vars]

qable(table1,markup='markdown',rgroup=rgroups)
```




```{r}
nsclc_data.trt= nsclc_data %>% filter(tx == 1);
nsclc_data.ctrl= nsclc_data %>% filter(tx == 0);
table2_summary.trt <- nsclc_data.trt %>%   select("obstime","survival.past.400") %>%  qsummary(.)
table2_summary.ctrl <- nsclc_data.ctrl %>%   select("obstime","survival.past.400") %>%  qsummary(.)

table2=list();
table2.trt=list();
table2.ctrl=list();
bin_list = c("europe",             "abnLDH","abnAlkphos")
for(i_var in 1:length(bin_list)){
(table2.trt[[i_var]] =nsclc_data.trt %>% group_by(.data[[bin_list[i_var]]]) %>% summary_table(., table2_summary.trt))
(table2.ctrl[[i_var]] =nsclc_data.ctrl %>% group_by(.data[[bin_list[i_var]]]) %>% summary_table(., table2_summary.ctrl))

table2[[i_var]] = cbind(table2.trt[[i_var]],table2.ctrl[[i_var]]);
}
rgroups2 =   rep(4,2)
names(rgroups2)=c("obstime","survival.past.400")
```

<span style="color:red">Note: </span>  Draw the Kaplan-Meier curve with no adjusted variables [Link](https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/).

```{r}
library(ggplot2)
library(survival)
library(ggfortify)
km_fit <- survfit(Surv(obstime, death) ~ tx, data=nsclc_data)

autoplot(km_fit)
```

```{r}
qable(table2[[1]],markup="markdown",rgroup=rgroups2)
```


```{r}
qable(table2[[2]],markup="markdown",rgroup=rgroups2)
```


```{r}
qable(table2[[3]],markup="markdown",rgroup=rgroups2)
```

## Inferential Analysis

```{r, cache = TRUE, message=FALSE}
library(lmtest)
library(sandwich)

nsclc_data.lm=lm(obstime~tx+europe+age+male+advdis+response+abnLDH+abnAlkphos+as.factor(ECOG),data=nsclc_data);

nsclc_data.robust_lm=coeftest(nsclc_data.lm, vcov = vcovHC(nsclc_data.lm, type="HC1"))

tmp=matrix(nsclc_data.robust_lm,nrow=dim(nsclc_data.robust_lm)[1],ncol=dim(nsclc_data.robust_lm)[2])
dimnames(tmp)=dimnames(nsclc_data.robust_lm);

caption_text = paste('Table 3: Results of multiple linear regression', sep='')
knitr::kable(
 tmp,  caption = caption_text
)
```



```{r,message=FALSE}
### Fit a Cox proportional hazard model

nsclc_data.cox <- coxph(Surv(obstime, death) ~tx+europe+age+male+advdis+response+abnLDH+abnAlkphos+as.factor(ECOG), data = nsclc_data)

sum_cox=summary(nsclc_data.cox)

caption_text = paste('Table 4: Results of Cox proportional hazard model', sep='')
knitr::kable(
 sum_cox$coef,  caption = caption_text
)
```

Check if adding more interactions will change results.

```{r}

nsclc_data.cox_int <- coxph(Surv(obstime, death) ~tx*(europe+age+male+advdis+response+abnLDH+abnAlkphos)+as.factor(ECOG), data = nsclc_data)

sum_cox_int=summary(nsclc_data.cox_int)

caption_text = paste('Table 5: Results of Cox proportional hazard model with interaction terms', sep='')
knitr::kable(
 sum_cox_int$coef,  caption = caption_text
)

```
```{r}
lklh_ratio=anova(nsclc_data.cox_int,nsclc_data.cox)

```

The likelihood ratio test for the two models yields a p-value of `r lklh_ratio[[4]][2]`.


## Sensitivity analysis


<span style="color:red">Note: </span>  Show the results of the sensitivity analysis in this section. You can also discuss any new analysis that are not in the interim report. 



# Discussion

<span style="color:red">Note: </span> Discuss the limitations of the presented projects, and comment on how this project enlightens future research or analysis.

# Session information
```{r}
print(sessionInfo(), local = FALSE)
```
