---
title: "Assessing Relationship Among Brain Regions in Mice"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
  pdf_document: default
---

<style type="text/css">
body{ /* Normal  */
      font-size: 14px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```


Team ID: 5

Name (tasks): Oscar Alvardo

Name (tasks): Karshima Johnson

Name (tasks): Christina Thai

Auditor (tasks):

***

## Introduction
This document is the final report for the Spring 2019 course STA 141A at University of California, Davis. It seeks to examine whether activity in one region affects the activity in another region. 

To reduce the variability of cortical responses to sensory stimuli, analyzing the neural spikes in multiple brain regions during spontaneous activity will yield more accurate, less noisy, results. 

Using the data gathered from contributors Nick Steinmetz, Marius Pachitariu, Carsen Stringer, Matteo Carandini, and Kenneth Harris' experiment, we seek to analyze the areas of the brain that are used in conjunction, regardless of stimuli. The original experiment investigates fluctuations in arousal levels indicated by running, pupil area, and whisking. Our variable of interest is the cortical responses as these behaviors are ongoing. 



## Background
```{r, message = FALSE, include=FALSE}
library("R.matlab") # Install 'R.matlab' if this is the first time you call it.
# Below VV Will vary depending on the local pathway for each computer
ephysroot = '/Users/christinathai/Desktop/R_programming/STA_141A/Project/'; #path of the data set 
mstr = 'Krebs'; # mouse names
### Load saved data:
exp_data<- readMat(paste(ephysroot,mstr,'_reduced.mat',sep=''));
# Check out data structure
str(exp_data)
```


In this project, we analyze the reduced data set `Reduced_Krebs.mat` containing Neuropixel recordings of a mouse named Krebs during spontaneous activity. The full data set can be found at this [webpage](https://janelia.figshare.com/articles/Eight-probe_Neuropixels_recordings_during_spontaneous_behaviors/7739750).

The data used has already been preprocessed using a Matlab script, which can be found in the code Appendix of the report. 


#Summary of Data
This report focuses on data collected from an experiment done with a mouse named Krebs. The neuron activation recorded from stimuli from whisking (measured by summed videographic motion energy within the whisker region).

This data set contains [spike trains](https://en.wikipedia.org/wiki/Neural_coding) of neurons in nine regions in the mouse brain.

The following represent elements of behavior data:
`avgframe` Average Frame of Recording <br />
`faceSVD` Singular Value Decompositon of Compressed Image of Face <br />
`motionMask` Corresponding Masks to Singular Values <brv/>

The data set being referenced `Reduced_Krebs.mat` contains the first 50 singular values and their masks. These are used to reproduce an image based on the average of the frames within the data set using infrared camera. 


The following columns represent the nine regions for Kreubs in the mouse brain:

`stall.CP` Caudate Putamen <br />
`stall.FrMoCtx` Frontal Motor Cortex<br />
`stall.HPF` Hippocampal Formation<br />
`stall.LS` Lateral Septum<br />
`stall.MB` Midbrain<br />
`stall.SC` Superior Colliculus<br />
`stall.SomMoCtxx` Somatomotor Cortex<br />
`stall.TH` Thalamus<br />
`stall.V1` Primary Visual Cortex <br />

The data set contains spike trains of neurons in nine regions of the mouse brain. Each corresponding column represents a specific region of the brain; for example, the vector `stall.CP` relates to the activity and activation of neurons located in the Caudate Putamen.

A more in depth explanation on the structure of our data is that it is a binary matrix with values of 1 denoting an existance of neural activity and 0 being that there is no neural activity at the specific time frame. Each row represents a neuron within the region. 


## Visual of Brain Regions
```{r, fig.width=50, fig.height=50,echo=FALSE}
#Image for the regions of the brain
library(png)
library(grid)
img <- readPNG("/Users/christinathai/Desktop/R_programming/STA_141A/Project/mouse_brain.png")
 grid.raster(img)
```
The figure above is the reconstructed probe locations of recordings in Krebs. The poles are generated using the borders of the brain regions and map depth of the region in microns relative to position of the principal axes of the body (AP/DV/LR). 

**Note:** AP refers the antero-posterior, DV refers to the dorso-ventral, and LR refers to the left-right axis formation.


```{r, echo = FALSE, eval = TRUE, message = FALSE}
brain_regions= ls(exp_data)[-c(1:3)]; 
n_time= dim(exp_data$faceSVD)[2];
n_sv = dim(exp_data$faceSVD)[1];
```

## Statistical questions of interest

To answer the primary scientific question of interest, we wish to assess the pairwise relationships between different brain regions within one subject. 


# Analysis Plan

## Population and study design

The initial study monitored large populations of neurons in awake head-fixed mice. These mice were unengaged from behavioral task and then their spontaneously performed behaviors (i.e. whisking, sniffing, and other facial movements) were monitored videographically. The study recorded simultaneous neurological data of six mice over nine sessions and worked to identify the relationship between these behaviors and their neural activity.  

Further, the studies were testing in various settings such as with and without darkness and recordings in darkness or with a gray screen. They observed no differences between these recordings, thus we take it to imply that Krebs' sample data is representative of the mouse's overall behavior in this experiment[.](https://www.biorxiv.org/content/biorxiv/early/2018/04/22/306019.full.pdf) 

In addition, we acknowledge that mouse models for genetic research prove to be excellent tools for insight into immune, endocrine, cardiovascular, skeletal, and in particular, nervous systems shared by mammals[.](https://www.genome.gov/10005834/background-on-mouse-as-a-model-organism) Thus conclusions obtained through the study may become applicable to humans.  




```{r}
spikes <- list()
spike_trains <- list()
#Code for extracting the spike trains
getspike <- function(i_region, t1, t2) {
  
    time_range = c(t1,t2)
      
  ### Extract the spike train for this region
  spikes_this_region = exp_data[[brain_regions[i_region]]]
  n_neurons = dim(spikes_this_region)[1]; 
  
  for(i_neuron in 1:n_neurons) {
        spk_times=which(spikes_this_region[i_neuron,time_range[1]:time_range[2]]>0);
        
        spikes[i_neuron] <- list(spk_times);
  }
  
  spike_trains[[i_region]] <- spikes[1:n_neurons]
  
  return(spike_trains)
}
for (i in 1:9) {
  spikes = getspike(i,0,n_time)
}
```

#### Spike Trains for Time Frame Indices 1000 to 2000

```{r echo = FALSE, eval = TRUE, message = FALSE}
plot_list = list()
for(i_region in 1:9) {
      time_range = c(1000, 2000); # indices of time frame
      
      ### Extract the spike train for this region
      spikes_this_region = exp_data[[brain_regions[i_region]]]
      
      ### Visualize the spike train 
      n_neurons = dim(spikes_this_region)[1]; 
      p = plot(x=0,y=0,pch=16,col="white",type="l",lwd=3,ylim=c(0,n_neurons),
               xlim=time_range,cex.lab=1,cex.main=1,ylab='Neuron',xlab='Time frame',
               main=paste0("Spike Train Region ", i_region, " Time ", time_range[1], 
                " to ", time_range[2]))
               #yaxt='n',xaxt='n',
      
      for(i_neuron in 1:n_neurons){
        spk_times =which(spikes_this_region[i_neuron,time_range[1]:time_range[2]]>0);
        if (length(spk_times)>0){
          points(y=rep(i_neuron,length(spk_times)),x=spk_times+time_range[1]-1,col="#000000",pch='.',cex=2)
        }
      }
      
      plot_list[[i_region]] = p
}
```


##Correlation Analysis
```{r, echo = FALSE, warning = FALSE}
library(dygraphs)

avg_neuron_list <- vector('list',9)
# par(mfrow = c(4,1))
for (i in 1:9) {
#Calculating the column sums 
spiketrain = exp_data[[brain_regions[i]]];
n_time = dim(spiketrain)[2]
n_neurons = dim(spiketrain)[1]
#Calculating the average number of neurons for each time frame
neurons_fired = colSums(spiketrain)
neurons.avg <- as.vector(neurons_fired/n_neurons)
avg_neuron_list[[i]] <-neurons.avg
# plot(as.ts(neurons.avg), xlab = "Time Frame", ylab = "Average number of neurons Fired")
}
```
## Overlap of Average Spike Trains for Each Region
Recall that the brain regions are:
`stall.CP` Caudate Putamen <br />
`stall.FrMoCtx` Frontal Motor Cortex<br />
`stall.HPF` Hippocampal Formation<br />
`stall.LS` Lateral Septum<br />
`stall.MB` Midbrain<br />
`stall.SC` Superior Colliculus<br />
`stall.SomMoCtxx` Somatomotor Cortex<br />
`stall.TH` Thalamus<br />
`stall.V1` Primary Visual Cortex <br />
```{r}
spike_data_frame = data.frame(time = (1:n_time), CP = avg_neuron_list[[1]], FrMoCtx = avg_neuron_list[[2]], HPF = avg_neuron_list[[3]], LS = avg_neuron_list[[4]], MB = avg_neuron_list[[5]], SC = avg_neuron_list[[6]], SoMoCtxx = avg_neuron_list[[7]], TH = avg_neuron_list[[8]], V1 =  avg_neuron_list[[9]])
dygraph(spike_data_frame) %>%
  dyOptions(labelsUTC = TRUE, fillGraph=TRUE, fillAlpha=0.1, drawGrid = FALSE, colors=c("#a987b7", "#59bad8", "#D8AE5A","#9b2b2b", "#8fb787", "#096010", "#0c62b7", "#f4b7f1", "#ff0008")) %>%
  dyRangeSelector() %>%
  dyCrosshair(direction = "vertical") %>%
  dyHighlight(highlightCircleSize = 8, highlightSeriesBackgroundAlpha = 0.2, hideOnMouseOut = FALSE)  %>%
  dyRoller(rollPeriod = 300)
```



From the above plot, we can see that the time from of most activity is between 20000 and 30000ms. Thus, we adjust our interval accordingly to analyze correlation.

We choose to examine correlation using the Kendall method. 

```{r}
library(ggcorrplot)

#Time frame of interest
time_frame <- c(20000:25000)


for (i in 1:9) {
  spikes[[i]] <- avg_neuron_list[[i]][time_frame]
}

spikes.cor.data = data.frame(CP = spikes[[1]], FrMoCtx = spikes[[2]], HPF = spikes[[3]], LS = spikes[[4]], MB = spikes[[5]], SC = spikes[[6]], SoMoCtxx = spikes[[7]], TH = spikes[[8]], V1 =  spikes[[9]])
spikes.cor <- cor(spikes.cor.data, method = "kendall")

ggcorrplot(spikes.cor)

```


```{r}
library(corrplot)
corrplot(spikes.cor, method="number", type = "upper")
```



#Results


# Discussion




#References

http://dev.biologists.org/content/133/11/2095



# Session information
```{r}
print(sessionInfo(), local = FALSE)
```
