---
title: "Assessing Regulatory Relationship Between Brain Regions during Spontaneous Activity: Interim Report"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
  pdf_document: default
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```


Team ID: 5

Name (tasks): Oscar Alvarado (Correlation Between Brain Regions)

Name (tasks): Karishma Johnson (Spike Trains, Krebs' Visual)

Name (tasks): Christina Thai (Summary of data, Generated Mouse Brain .png, ..)


***



# Introduction
This document is the interim report for STA 141A, Spring 2019. It seeks to examine whether their exists a regulatory relationship among regions. In other words, we will be investigating if activity in one region affects the activity in another region. 

To reduce the variability of cortical responses to sensory stimuli, analyzing the neural spikes in multiple brain regions during spontaneous activity will yield more accurate, less noisy, results. 

Using the data gathered from contributors Nick Steinmetz, Marius Pachitariu, Carsen Stringer, Matteo Carandini, and Kenneth Harris' experiment, we seek to analyze the areas of the brain that are used in conjunction, regardless of stimuli. The original experiment investigates fluctuations in arousal levels indicated by running, pupil area, and whisking. Our variable of interest is the cortical responses as these behaviors are ongoing. 



## Background
```{r, message = FALSE, include=FALSE}
library("R.matlab") # Install 'R.matlab' if this is the first time you call it.

# Below VV Will vary depending on the local pathway for each computer
ephysroot = '/Users/christinathai/Desktop/R_programming/STA_141A/Project/'; #path of the data set 

mstr = 'Krebs'; # mouse names

### Load saved data:
exp_data<- readMat(paste(ephysroot,mstr,'_reduced.mat',sep=''));

# Check out data structure
str(exp_data)
```


In this project, we analyze the reduced data set `Reduced_Krebs.mat` containing Neuropixel recordings of a mouse named Krebs during spontaneous activity. The full data set can be found at this [webpage](https://janelia.figshare.com/articles/Eight-probe_Neuropixels_recordings_during_spontaneous_behaviors/7739750).

The data used has already been preprocessed using a Matlab script, which can be found in the code Appendix of the report. 


#Summary of Data
This report focuses on data collected from an experiment done with a mouse named Krebs. The neuron activation recorded from stimuli from whisking (measured by summed videographic motion energy within the whisker region).

This data set contains [spike trains](https://en.wikipedia.org/wiki/Neural_coding) of neurons in nine regions in the mouse brain.

The following represent elements of behavior data:
`avgframe` Average Frame of Recording <br />
`faceSVD` Singular Value Decompositon of Compressed Image of Face <br />
`motionMask` Corresponding Masks to Singular Values <brv/>

The data set being referenced `Reduced_Krebs.mat` contains the first 50 singular values and their masks. These are used to reproduce an image based on the average of the frames within the data set using infrared camera. 


The following columns represent the nine regions for Kreubs in the mouse brain:

`stall.CP` Caudate Putamen <br />
`stall.FrMoCtx` Frontal Motor Cortex<br />
`stall.HPF` Hippocampal Formation<br />
`stall.LS` Lateral Septum<br />
`stall.MB` Midbrain<br />
`stall.SC` Superior Colliculus<br />
`stall.SomMoCtxx` Somatomotor Cortex<br />
`stall.TH` Thalamus<br />
`stall.V1` Primary Visual Cortex <br />

The data set contains spike trains of neurons in nine regions of the mouse brain. Each corresponding column represents a specific region of the brain; for example, the vector `stall.CP` relates to the activity and activation of neurons located in the Caudate Putamen.


## Visual of Brain Regions
```{r, fig.width=50, fig.height=50,echo=FALSE}
#Image for the regions of the brain
library(png)
library(grid)
img <- readPNG("/Users/christinathai/Desktop/R_programming/STA_141A/Project/mouse_brain.png")
 grid.raster(img)
```
The figure above is the reconstructed probe locations of recordings in Krebs. The poles are generated using the borders of the brain regions and map depth of the region in microns relative to position of the principal axes of the body (AP/DV/LR). 

**Note:** AP refers the antero-posterior, DV refers to the dorso-ventral, and LR refers to the left-right axis formation.


```{r, echo = FALSE, eval = TRUE, message = FALSE}
brain_regions= ls(exp_data)[-c(1:3)]; 
n_time= dim(exp_data$faceSVD)[2];
n_sv = dim(exp_data$faceSVD)[1];
```

## Statistical questions of interest

To answer the primary scientific question of interest, we wish to assess the pairwise relationships between different brain regions within one subject. 



# Analysis Plan
Since our sample size (number of mice in experiment) is extremely small, n = 1, we cpropose to use Kendall Tau Correlation method since it will outperform Pearson and Spearman's. 


Additionally, Kendall's Tau is ideal because it measures the extent to which, as one variable increases, the other variable tends to increase, without requiring the increase to be represented by a linear relationship.

Our goal in this project is to establish if there exists a regulatory relationship among the nine brain regions by
analyzing the neural spike trains in the nine brain regions and determining if there is a correlation between them. We will do this 
by computing the Spike Time Tiling Coefficient(STTC), which is a method that was proposed by Catherine S. Cutts and Stephen J. Eglen in the paper "Detecting Pairwise Correlations in Spike Trains: An Objective Comparison of Methods and Application to the Study of Retinal Waves" in the Journal of Neuroscience. STTC is defined as STTC = 0.5((P_A-T_B)/(1-P_A*T_B) + (P_B-T_A)/(1-P_B*T_A)) where 
P_A = the proportion of spike from A which lie within +/- delta t of any spike from B, P_B = similar as P_A
T_A = the proportion of total recording time which lies within +/- delta t of any spike from A, T_B = similar as T_B
In their paper, Cutts and Eglen argue that STTC is superior to other coefficient measurements, particularly the 
coefficient index, which is a factor by which the firing rate of neuron A increases over its mean value if measured within a fixed window of spikes from neuron B. Unlike the coefficient index, STTC is not bounded above nor is it cofounded by the firing rate. 
In fact, STTC is bounded between values -1 and 1, -1 indicating strong anti-correlation, 0 no correlation, and 1 indicating identical spike trains. Also, the fact that STTC is not confounded by the firing rate indicates that it is a much fairer way of comparing correlations. As stated in Cutts and Watts paper,
"Firing rates frequently vary across age, phenotype, and presence of pharmacological agonists and so this calls into question the results of many correlation analyses: some
conclusions about differences in correlations may be due to the confounding effect of the firing rates, and not the correlations themselves." 
In essence, we will take a spike train from one of the nine brain regions, and compute the STTC of it with the spike trains of the remaining 8 brain regions, and repeat this for all brain regions. We will achieve this by using the R package meaRtools, which has the algorithm for computing STTC.  

In addition, we will use the STTC to produce a correlogram. Suppose we have two spike trains A and B, each being from a different brain region. For a given arbitrary value tau, we will cross correlate A and B by adding tau to the spike time of B. This will help detect correlations between the brain regions when temporal delays occur. Furthermore, we will create a lattice plot to demonstrate the brain regions pairwise correlations. By creating the lattice plot, we will be able to see clearly which regions have more correlation with one another.   



## Population and study design

The initial study monitored large populations of neurons in awake head-fixed mice. These mice were unengaged from behavioral task and then their spontaneously performed behaviors (i.e. whisking, sniffing, and other facial movements) were monitored videographically. The study recorded simultaneous neurological data of six mice over nine sessions and worked to identify the relationship between these behaviors and their neural activity.  

Further, the studies were testing in various settings such as with and without darkness and recordings in darkness or with a gray screen. They observed no differences between these recordings, thus we take it to imply that Krebs' sample data is representative of the mouse's overall behavior in this experiment[.](https://www.biorxiv.org/content/biorxiv/early/2018/04/22/306019.full.pdf) 

In addition, we acknowledge that mouse models for genetic research prove to be excellent tools for insight into immune, endocrine, cardiovascular, skeletal, and in particular, nervous systems shared by mammals[.](https://www.genome.gov/10005834/background-on-mouse-as-a-model-organism) Thus conclusions obtained through the study may become applicable to humans.  


# Results

## Spike Trains 

Spike trains are a mode of information transmission within the nervous system and serve as the basis of our study design[.](https://neuronaldynamics.epfl.ch/online/Ch7.S1.html) They are formed by a sequence of recorded times where neurons are firing action potentional. These firings are measured in terms of milliseconds and form a spike. Sequences of spikes are referred to as spike trains[.](http://www.stat.cmu.edu/~kass/papers/SpikeTrain.pdf) Neuron firing is induced by physical sensory stimuli. In the case of the original study, this is in correspondence with whisking, pupil area, and running. "Spike trains often last as long as the stimulus is present[.](http://www.neuwritewest.org/blog/2015/1/3/ask-a-neuroscientist-whats-a-spike-train)" Thus, when looking at the time indices 100 to 1000 there are notable blank spaces where spikes are not present. These could potentially be attributed to a lack of stimulus within the region, but we consider that since it is near the beginning of the data recording period, it may have been an issue in the recording of data itself over a lack of neurological activity.

#### Time Frame Indices 100 to 1000
```{r echo = FALSE,eval = TRUE, message = FALSE}
library(R.matlab) 
library(ggplot2)

plot_list = list()
for(i_region in 1:9) {
      time_range = c(100, 1000); # indices of time frame
      
      ### Extract the spike train for this region
      spikes_this_region = exp_data[[brain_regions[i_region]]]
      
      ### Visualize the spike train 
      n_neurons = dim(spikes_this_region)[1]; 
      p = plot(x=0,y=0,pch=16,col="white",type="l",lwd=3,ylim=c(0,n_neurons),
               xlim=time_range,cex.lab=1,cex.main=1,ylab='Neuron',xlab='Time frame',
               main=paste0("Spike Train Region ", i_region, " Time ", time_range[1], 
                " to ", time_range[2]))
               #yaxt='n',xaxt='n',
      
      for(i_neuron in 1:n_neurons){
        spk_times =which(spikes_this_region[i_neuron,time_range[1]:time_range[2]]>0);
        if (length(spk_times)>0){
          points(y=rep(i_neuron,length(spk_times)),x=spk_times+time_range[1]-1,col="#000000",pch='.',cex=2)
        }
      }
      
      plot_list[[i_region]] = p

}
```

#### Spike Trains for Time Frame Indices 1000 to 2000

```{r echo = FALSE, eval = TRUE, message = FALSE}
plot_list = list()
for(i_region in 1:9) {
      time_range = c(1000, 2000); # indices of time frame
      
      ### Extract the spike train for this region
      spikes_this_region = exp_data[[brain_regions[i_region]]]
      
      ### Visualize the spike train 
      n_neurons = dim(spikes_this_region)[1]; 
      p = plot(x=0,y=0,pch=16,col="white",type="l",lwd=3,ylim=c(0,n_neurons),
               xlim=time_range,cex.lab=1,cex.main=1,ylab='Neuron',xlab='Time frame',
               main=paste0("Spike Train Region ", i_region, " Time ", time_range[1], 
                " to ", time_range[2]))
               #yaxt='n',xaxt='n',
      
      for(i_neuron in 1:n_neurons){
        spk_times =which(spikes_this_region[i_neuron,time_range[1]:time_range[2]]>0);
        if (length(spk_times)>0){
          points(y=rep(i_neuron,length(spk_times)),x=spk_times+time_range[1]-1,col="#000000",pch='.',cex=2)
        }
      }
      
      plot_list[[i_region]] = p

}
```



## Sensitivity analysis
 - the study of how the uncertainity in the output of a mathemadical model or system can be divided and allocated to different soures of uncertainity in its inputs



# Future analysis plan 
Since we have analyzed the correlation for each brain region. Another insightful feature to our report would be to create a prediction model to examine the strength of the relationship of spontaneous activity on neural activity. Additionally, incorporating more stimuli can give us a better understanding of how which stimuli triggers activity in specific regions of the brain.

#References

http://dev.biologists.org/content/133/11/2095



# Session information
```{r}
print(sessionInfo(), local = FALSE)
```
